---
title: "Big Data Analytics"
subtitle: "Assignment 1"
date: "26/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error = F,
                      message = F,
                      warning = F,
                      comment = "")
library(ggplot2)
library(dplyr)
library(reshape2)
```

Reading the data into R workspace

```{r}
diamond <- read.csv("DiamondData.csv")
```

# Task 1.

First printing the summary of all the variables in the dataset

```{r}
summary(diamond)
```

The errors sin the dataset includes:

1. The `carat` variable varies from a minimum value of 0.2 to maximum of 5.01
1. The `Very Good` level under `cut` attribute is also mistyped as `Very Geod`
1. NA's in all the attributes
1. Recalculating the value of `depth`

**Correcting the level of `cut` variable**

```{r}
diamond$cut[diamond$cut == "Very Geod"] <- "Very Good"
diamond$cut <- as.factor(as.character(diamond$cut))
summary(diamond$cut)
```

**Removing all the rows containing any NAs's in the dataset**

```{r}
diamond <- diamond[complete.cases(diamond), ]
dim(diamond)
```

**Correcting the range for `carat` variable**

```{r}
diamond <- diamond[diamond$carat >= 0.2 & diamond$carat <= 5.01,]
summary(diamond$carat)
```

```{r}
dim(diamond)
```

**Recalculating the values for `depth` variable**

```{r}
diamond$depth <- 2*diamond$z/(diamond$x+diamond$y)
diamond <- diamond[complete.cases(diamond), ]
```

# Task 2

```{r}
summary(diamond)
```

```{r}
ggplot(diamond, aes(x=carat,y=depth)) +
  geom_point(aes(col=cut)) +
  labs(
    x = "Carat",
    y = "Depth",
    title = "Depth Vs Carat colored by cut quality"
  )

ggplot(diamond, aes(x=carat,y=depth)) +
  geom_point(aes(col=color)) +
  labs(
    x = "Carat",
    y = "Depth",
    title = "Depth Vs Carat colored by diamond color"
  )

ggplot(diamond, aes(x=carat,y=depth)) +
  geom_point(aes(col=clarity)) +
  labs(
    x = "Carat",
    y = "Depth",
    title = "Depth Vs Carat colored by diamond clarity"
  )
ggplot(diamond, aes(x=carat,y=table)) +
  geom_point(aes(col=cut)) +
  labs(
    x = "Carat",
    y = "Width",
    title = "Width Vs Carat colored by cut quality"
  )

ggplot(diamond, aes(x=carat,y=table)) +
  geom_point(aes(col=color)) +
  labs(
    x = "Carot",
    y = "Width",
    title = "Width Vs Carat colored by diamond color"
  )

ggplot(diamond, aes(x=carat,y=table)) +
  geom_point(aes(col=clarity)) +
  labs(
    x = "Carat",
    y = "Depth",
    title = "Depth Vs Carat colored by diamond clarity"
  )
```

```{r}
ggplot(diamond, aes(x=carat, fill=color)) +
  geom_histogram() +
  labs(
    x = "Carat",
    y = "Frequency",
    title = "Carat Distribution colored by diamond clarity"
  ) + 
  scale_color_grey()

ggplot(diamond, aes(x=depth, fill=color)) +
  geom_histogram() +
  labs(
    x = "Depth",
    y = "Frequency",
    title = "Depth Distribution colored by diamond clarity"
  ) + 
  scale_color_grey()

ggplot(diamond, aes(x=table, fill=color)) +
  geom_histogram() +
  labs(
    x = "Widt",
    y = "Frequency",
    title = "Diamond Top Width Distribution colored by diamond clarity"
  ) + 
  scale_color_grey()
```

```{r}
diamond %>%
  group_by(color) %>%
  summarise(
    mean.depth = mean(depth)
  ) %>%
  ggplot(aes(x="", y=mean.depth, fill=color)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0)+
  theme_void() +
  labs(
    title = "Diamond mean Depth for different colors"
  )
```

```{r}
diamond %>%
  group_by(color) %>%
  summarise(
    mean.width = mean(table)
  ) %>%
  ggplot(aes(y=mean.width, x=color)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Diamond mean top width for different colors"
  ) + 
  coord_flip()
```

# Task 3

## Part (a)

The following results shows the summary statistics for the `price` variable

```{r}
summary(diamond$price)
```

Now plotting the histogram for `price` variable

```{r}
diamond %>%
  ggplot(aes(price)) +
  geom_histogram() +
  labs(
    title = "Price Distribution"
  )
```

From the above histrogram, we can observe that `price` variable is positively skewed. 

## Part (b)

```{r}
price.group <- cut(diamond$price, 3, include.lowest=TRUE, labels=c("Low", "Med", "High"))
table(price.group)
```

The following table shows the mea value of different numerical attributes in the dataset in different price groups

```{r}
diamond$price.group <- price.group
t(diamond %>%
    group_by(price.group) %>%
    summarise_all(mean))
```

## Part (c)

```{r}
diamond %>%
  ggplot(aes(x=cut,y=price)) +
  geom_boxplot() +
  labs(
    title = "Prices for different cut types"
  )
```

## Part (d)

```{r}
nums <- unlist(lapply(diamond, is.numeric))  
diamond.nums <- diamond[,nums]
cormat <- round(cor(diamond.nums),2)
melted_cormat <- melt(cormat)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  labs(
    x = "",
    y = "",
    title = "Heat Map/Correlation Matrix"
  )
```

From the above plot, we can see that `carat` `x` and `y` are 3 most correlated variables with price.

# Task 4

```{r}
cut <- diamond$cut
clarity <- diamond$clarity
table(cut, clarity)
```

```{r}
ggplot(diamond, aes(x=carat,y=price)) +
  geom_point(aes(col=cut)) +
  labs(
    x = "Carat",
    y = "Price",
    title = "Price Vs Carat colored by cut quality"
  )

ggplot(diamond, aes(x=carat,y=price)) +
  geom_point(aes(col=color)) +
  labs(
    x = "Carat",
    y = "Price",
    title = "Price Vs Carat colored by diamond color"
  )
```

# Task 5
  
## Part (a)

```{r}
diamond$volume <- diamond$x*diamond$y*diamond$z
ggplot(diamond, aes(x=volume,y=price)) +
  geom_point() +
  labs(
    x = "Volume",
    y = "Price",
    title = "Price Vs Volume"
  )
```


## Part (b)

```{r}
cor(diamond$carat, diamond$volume)
```

From the above correlatio cefficient, we can tell that `volume` and `carat` are highly correlated.

```{r}
ggplot(diamond, aes(x=volume,y=carat)) +
  geom_point() +
  labs(
    x = "Volume",
    y = "Carat",
    title = "Carat Vs Volume"
  ) + 
  geom_smooth(method='lm')
```

## Part (c)

```{r}
cor(diamond$table, diamond$depth)
```

From the above correlatio cefficient, we can tell that `table` and `depth` are not strongly correlated and the correlation is negative.

```{r}
ggplot(diamond, aes(x=depth,y=table)) +
  geom_point() +
  labs(
    x = "Depth",
    y = "Table",
    title = "Table Vs Depth"
  ) + 
  geom_smooth(method='lm')
```

## Part (d)

```{r}
cormat
```