---
output:
  html_document: default
  pdf_document: default
---


```{r}
library(readr)
```


#install.packages("rmarkdown")
#rmarkdown::render()

#Sys.sleep(30)

#install.packages("knitr")

#library(knitr)
#knitr::knit2html('Task - B.Rmd')

```{r}
DiamondDataComplete <- read.csv(file.choose(), header=TRUE) 
```


```{r}
View(DiamondDataComplete)
```


```{r}
summary(DiamondDataComplete)

#   carat            cut                color          
#Min.   :0.2000   Length:50000       Length:50000      
#1st Qu.:0.4000   Class :character   Class :character  
#Median :0.7000   Mode  :character   Mode  :character  
#Mean   :0.7974                                        
#3rd Qu.:1.0400                                        
#Max.   :5.0100                                        
#   clarity            depth           table           price      
#Length:50000       Min.   :43.00   Min.   :43.00   Min.   :  326  
#Class :character   1st Qu.:61.00   1st Qu.:56.00   1st Qu.:  949  
#Mode  :character   Median :61.80   Median :57.00   Median : 2401  
#Mean   :61.75   Mean   :57.45   Mean   : 3925  
#3rd Qu.:62.50   3rd Qu.:59.00   3rd Qu.: 5312  
#Max.   :79.00   Max.   :95.00   Max.   :18823  
#     x               y                z         
#Min.   : 0.00   Min.   : 0.000   Min.   : 0.000  
#1st Qu.: 4.71   1st Qu.: 4.720   1st Qu.: 2.910  
#Median : 5.70   Median : 5.710   Median : 3.520  
#Mean   : 5.73   Mean   : 5.732   Mean   : 3.538  
#3rd Qu.: 6.54   3rd Qu.: 6.540   3rd Qu.: 4.030  
#Max.   :10.74   Max.   :31.800   Max.   :31.800 
```


```{r}
attach(DiamondDataComplete)
```


```{R}
boxplot(DiamondDataComplete$price)
```


```{r}
hist(DiamondDataComplete$price) #Price is Right Skewed 
```


```{r}
hist(DiamondDataComplete$depth) #Depth is Normally Distributed
```




```{r}
hist(DiamondDataComplete$table) #Table is left skewed
```



```{R}
hist(DiamondDataComplete$x) #Length of the diamond is right skewed
```



```{r}
hist(DiamondDataComplete$y) #Width of the diamond is right skewed
```


```{r}
hist(DiamondDataComplete$z) #Depth of the diamond is left skewed 
```


```{r}
boxplot(DiamondDataComplete$z) #Complete Data is distributed left side
```




```{r}
names(DiamondDataComplete)

#[1] "carat"   "cut"     "color"   "clarity" "depth"   "table"  
#[7] "price"   "x"       "y"       "z
```


```{r}
table(cut)
#cut
#Fair      Good     Ideal   Premium Very Good 
#1500      4539     20011     12770     11180 
```


```{r}
table(color)


color
#D     E     F     G     H     I     J 
#6269  9097  8854 10463  7666  5029  2622
```




```{r}
table(clarity)

clarity
#I1    IF   SI1   SI2   VS1   VS2  VVS1  VVS2 
#679  1667 12110  8501  7582 11365  3404  4692
```



```{r}
DiamondDataComplete$cut = factor(DiamondDataComplete$cut, levels = c('Fair','Good','Ideal', 'Premium','Very Good'), labels = c(0,1,2,3,4))




DiamondDataComplete$color = factor(DiamondDataComplete$color, levels = c('D','E','F','G','H','I','J'), labels = c(0,1,2,3,4,5,6))





DiamondDataComplete$clarity = factor(DiamondDataComplete$clarity, levels = c('I1','IF','SI1','SI2','VS1','VS2','VVS1','VVS2'), labels = c(0,1,2,3,4,5,6,7))
```


```{r}
str(DiamondDataComplete$cut)




str(DiamondDataComplete$color)




str(DiamondDataComplete$clarity)

```


```{r}
model <- lm(price ~ cut + color + clarity + x + y + z + table + depth + carat)
```


```{r}
summary(model) #All are significant except depth(z) 

#Multiple R-squared:  0.9194,	Adjusted R-squared:  0.9193 
```


```{r}
#Now finding the correlation between the features
cor(price, carat)  #highly correlated = 0.9213562



cor(price, table)  #0.1278579



cor(price, x) #0.884927



cor(price, y) #0.8817025




cor(price, z) #0.8607349



cor(price, depth) #-0.01213396 - Negatively correlated
```






```{r}
require(caTools) 




library(caTools)
```




```{r}
#Data is Divided into Training and testing Data 





data_split <- sort(sample(nrow(DiamondDataComplete), nrow(DiamondDataComplete)*.75))





train_Data <- DiamondDataComplete[data_split, ]



test_data <- DiamondDataComplete[-data_split, ]





nrow(train_Data)  #37500




nrow(test_data)   #12500
```





```{r}
#Model built on Training data
train_model <- lm(train_Data$price ~., data = train_Data)
```



```{r}
summary(train_model)
```



```{r}
#y - width and z - depth are not significant, hence removed
train_model_final <- lm(train_Data$price ~ train_Data$carat + train_Data$cut + train_Data$color + train_Data$clarity + train_Data$depth + train_Data$depth + train_Data$table + train_Data$x, data = train_Data)






summary(train_model_final)  #All are significant 

#Multiple R-squared:  0.9191,	Adjusted R-squared:  0.919
```


```{r}
#predicting the test data and findind RMSE
#new_Data <- data.frame(price_test = test_Data)
#output <- cbind(testdata, prediction)

#new_test <- data.frame(testData = test_data$price)
#prediction <- predict(train_model_final, newdata = new_test, interval = "confidence")
#summary(prediction)
#nrow(prediction)

#model bulding on test data




test_model <- lm(test_data$price ~., data = test_data)





test_model





summary(test_model)
```



```{r}
#y - width and z - depth are not significant, hence removed






test_model_final <- lm(test_data$price ~ test_data$carat + test_data$cut + test_data$color + test_data$clarity + test_data$depth + test_data$table + test_data$price + test_data$x, data = test_data)




summary(test_model_final)
```



```{r}
pred_train <- predict(train_model_final)
```



```{r}

pred_test <- predict(test_model_final)
```

```{r}
error = test_data$price - pred_test
```

```{r}
error
```

```{r}
summary(error)
```



```{r}
rmse_test <- sqrt(mean(error^2))
```


```{r}
rmse_test #1141.406 - RMSE of test data
```


```{r}
actual_pred <- data.frame(cbind(test_data$price, pred_test))
```






```{r}
correlation_accuracy <- cor(actual_pred)
```

```{r}
correlation_accuracy
#              V1    pred_test
#V1        1.0000000 0.9576256
#pred_test 0.9576256 1.0000000    #Correlation accuracy of test data is = 95.7%
```



```{r}

head(actual_pred)
#V1 pred_test
#1   957  1948.010
#2  4554  5138.437
#3 13853 10588.046
#4  4858  4796.315
#5 15814 14239.720
#6  2431  3167.320
```




```{r}
#predict the train model and find RMSE
prediction_train <- predict(train_model_final)
```




```{r}
error_train = train_Data$price - prediction_train
```





```{r}
error_train
```


```{r}

rmse_train <- sqrt(mean(error_train^2))

```


```{r}
rmse_train  #1126.223 - RMSE of train data
```




```{r}
plot(test_data$price, pred_test)
```




```{r}
plot(train_Data$price, pred_train)
```




```{r}
#Normalization of the diamond data






normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

```



```{r}
#Applying normalization on only numerical data
DiamondDataComplete_norm <- normalize(DiamondDataComplete[c(1,5,6,7,8,9,10)])
DiamondDataComplete_norm
View(DiamondDataComplete_norm)  #Data is normalized and removed categorical variables
```





```{r}
#Building Multiple linear regression model on normalized diamond data

norm_model <- lm(DiamondDataComplete_norm$price ~., data = DiamondDataComplete_norm)
norm_model

```



```{R}
summary(norm_model) #Z(depth) is not significant
```


```{r}
#Removing the insignificant and building the model again





final_norm_model <- lm(DiamondDataComplete_norm$price ~ DiamondDataComplete_norm$carat + DiamondDataComplete_norm$depth + DiamondDataComplete_norm$table + DiamondDataComplete_norm$x + DiamondDataComplete_norm$y, data = DiamondDataComplete_norm)
final_norm_model





summary(final_norm_model)

#Multiple R-squared:  0.8588,	Adjusted R-squared:  0.8588 

```




```{r}
#Splitting data in to train and test

data_split <- sort(sample(nrow(DiamondDataComplete_norm), nrow(DiamondDataComplete_norm)*.75))


train_Data_norm <- DiamondDataComplete_norm[data_split, ]
test_Data_norm <- DiamondDataComplete_norm[-data_split, ]
```



```{r}
#Bulding model on train data

train_norm <- lm(train_Data_norm$price ~ train_Data_norm$carat + train_Data_norm$depth + train_Data_norm$table + train_Data_norm$x + train_Data_norm$y + train_Data_norm$z)
train_norm
summary(train_norm)

```




```{r}
#Z(depth) is not significant
#Removing the z - depth and building final train model

train_norm_final <- lm(train_Data_norm$price ~ train_Data_norm$carat + train_Data_norm$depth + train_Data_norm$table + train_Data_norm$x + train_Data_norm$y, data = train_Data_norm)
train_norm_final
summary(train_norm_final) #All are significant
#Multiple R-squared:  0.8573,	Adjusted R-squared:  0.8573
```


```{r}
#predicting the test data using train model

test_norm <- lm(test_Data_norm$price ~., data = test_Data_norm)
summary(test_norm)  #All are significant

pred_test_norm <- predict(test_norm)
pred_train_norm <- predict(train_norm_final)


error_train_norm <- train_Data_norm$price - pred_train_norm
error_test_norm <- test_Data_norm$price - pred_test_norm

error_test_norm
error_train_norm
```

```{r}
rmse_test_norm <- sqrt(mean(error_test_norm^2))
rmse_test_norm #0.07909287 - RMSE of test data 


rmse_train_norm <- sqrt(mean(error_train_norm^2))
rmse_train_norm #0.07942949 - RMSE of train data

plot(train_Data_norm$price, pred_train_norm)
plot(test_Data_norm$price, pred_test_norm)

```

```{r}

prediction_test_norm <- predict(test_norm) 
actual_pred1 <- data.frame(cbind(test_Data_norm$price, pred_test_norm))
correlation_accuracy1 <-  cor(actual_pred1)
correlation_accuracy1

```


Differences in both the models

1st model - The Multiple Linear regression is built on not normalized data with multiple variables. The data is splitted into train and test data as given in the assignment and performed analysis nd built the model on complete Diamond data. 

Findings of 1st model - 
RMSE - 1141.406 - RMSE of test data
Correlation accuracy of test data is = 95.7%




2nd Model - This model is standardized so, the data will be standardized between 0 and 1, so, the RMSE also gives its results
between 0 and 1. 
Findings of 2nd model - 

RMSE of training data - 0.07882821
RMSE of testing data  - 0.07963757
Correlation Accuracy of normalized data - 92.7%











